Scalar and Representative Observables, and Their Associated Statistical Model, for Polarimetric SAR Data

### 2014-03-10 paper_dev/article.tex
integrate Ian's changes

%\author{Thanh-Hai Le, Ian McLoughlin}
\author{Thanh-Hai~Le,
        Ian~McLoughlin, 
	and Chan-Hua~Vun%
\thanks{Thanh-Hai~Le and Chan-Hua~Vun are with School of Computer Engineering, 
Nanyang Technological University, Singapore. Ian~McLoughlin is with the School of Information Science and Technology,
University of Science and Technology of China.
}% <-this % stops a space
%\thanks{The authors wish to thank Dr. Ken-Yoong Lee and Dr. Timo Brestchneider of EADS Innovation-Works Singapore for 
%	providing us the RADAR-SAT2 imagery used in this paper. }% <-this % stops a space
\thanks{Manuscript received ?, 2014; revised ?.}}

\markboth{IEEE Journal on Selected Topic in Applied Earth Observation and Remote Sensing,~Vol.~?, No.~?, ?~2014}%
{ Le \MakeLowercase{\textit{et al.}}:  Scalar and Representative Observables, and Their Associated Statistical Model, for Polarimetric SAR Data}

@@

%How do others solve this problem?
%The complex POLSAR data have been statistically modelled as following the complex Wishart distribution,
%  which apparently is multidimensional, complex and thus not very intuitive.
%There has been a few scalar observables with accompanying statistical models proposal for POLSAR
A few scalar POLSAR observables with accompanying statistical models have been proposed \cite{Conradsen_2003_TGRS_4, Alberga_2008_IJRS_4129, Joughin_1994_TGRS_562, Lee_1994_TGRS_1017, Touzi_1996_TGRS_519, Lopez-Martinez_2003_TGRS_2232, Erten_2012_Sensors_2766},
  but none of them is able to provide widely-accepted scalar discrimination measures.
%  but none of them has led to scalar discrimination measures proposal.
As such no observable can be considered as being highly representative of this multi-dimensional data,
  which severely limits their applicability in practical data processing applications.
Alternatively, a few popular POLSAR discrimination measures are reviewed \cite{Lee_1994_IJRS_2299, Anfinsen_2007_ESA_POLINSAR, Kersten_2005_TGRS_519, Lee_2011_IGARSS_3740, Cao_2007_TGRS_3454},
%From an alternative approach, a few POLSAR discrimination measures have been proposed,
  but all of them are shown to be based on the likelihood ratio statistics.
This statistical test should ideally be based on an exact and consistent distribution
   but so far only asymptotic distributions have been demonstrated \cite{Conradsen_2003_TGRS_4}.  
%This statistical test should be based on an exact and consistent statistical distribution
%  but so far has only shown to be based on an asymptotic distribution.

%Research Objective: Give your precise problem statement
This paper presents scalar and representative observables,
   and one associated generic statistical model,
   that describe multi-dimensional POLSAR data and provide a consistent foundation for the derivation of discrimination measures.   
%This article hence presents a scalar and representative observable, and its associated generic statistical model to describe multi-dimensional POLSAR data and provide a consistent foundation for the derivation of POLSAR discrimination measures.
%Now State your approach
The development is based on the common assumption that POLSAR data follows the complex Wishart distribution.  
Subsquently, the generic statistical model for the covariance matrix determinant is presented as being just a scalar projection of the multidimensional data.
This model is then used to derive several scalar and consistent statistical distributions suggesting that their associated observables are capable of being used as discrimination measures for POLSAR.
%This study started off with the observation that the determinant of the covariance matrix is widely used in POLSAR discrimination measures.
%However, our survey of existing scalar models for POLSAR indicates that no statistical model for this observable has been published so far.
%Consequently, in this article, the statistical behaviour of the determinant of POLSAR covariance matrix is further investigated.
%Prof Vuns: comment: the article provides the result, not the journey!
%POLSAR data have been statistically modelled as following the complex Wishart distribution.
%Consequently, the generic statistical model for the covariance matrix determinant is presented as being just a scalar projection of the multidimensional data.
%This model is then used to derive several scalar and consistent statistical descriptions suggesting that their associated observables are capable of being used as discrimination measures for POLSAR data.
%
%***IVM: the following sentence doesn't make sense, also, I don't think it's good to talk about 'models of a model'.  Maybe 'model variants' instead?
%%%***IVM***The specific two and three dimensional models of these models, which %also include the derived discrimination measures, are originally designed against both partial and full polarimetric SAR data.

This paper also shows that %It is also shown that
   the specific one dimensional (1-D) version of the proposed  model 
   matches %perfectly with 
   the traditional statistical model used for SAR intensity.
This effectively incorporates the existing SAR theory under the umbrella of the proposed scalar approach for multi dimensional POLSAR.
The different discrimination proposals for SAR and POLSAR are reviewed in light of this and the  proposal in this paper is shown to provide a strong, unifying and consistent foundation. % relating them all together.
The applicability of these theoretical models will be illustrated by experiments where the specific 1, 2 and 3-D versions of the proposed models are validated against real-life captured data.    
The other beneficial implication of the proposed approach will also be demonstrated,
  where the proposed theore is applied, as an example, to evaluate POLSAR speckle filters.

@@

%Article Organization: Show that you deliver on the objectives and hence solve the stated problem
The remainder of this article is organized as follows.
After the next section reviews existing discrimination measures as well as scalar observables models for POLSAR, section \ref{sec:theoretical_model}
  derives the generic statistical model for the determinant of the POLSAR covariance matrix
  and proposes several new discrimination measures for the multi-dimensional data.
After section \ref{sec:sar_special_case_of_polsar} illustrates the match between the specific model of one dimensional and the traditional model for SAR intensity,
  section \ref{sec:link_sar_polsar} links the disparate discrimination measure proposals for the traditional SAR and the more recent POLSAR data.
%Section \ref{sec:polsar_models_validation} then validates the specific one, two and three dimensional versions of the proposed model against real-life captured data.
The applicability of these models against practical data is illustrated in section \ref{sec:polsar_models_validation}.
%Section \ref{sec:discussion} presents a high-level discussion of the results presented before
Then section \ref{sec:application} presents a simple demonstration where an existing technique of SAR processing is extended for POLSAR data before
  section \ref{sec:conclusion} finally concludes the paper.

@@

%This section outlines related publication.
Section \ref{sec:lit_models} reviews various statistical models used for different scalar observables for POLSAR.
  where it is shown that none are able to provide statistically consistent discrimination measures. 
Section \ref{sec:lit_measures} further strengthens those findings by discussing discrimination measures that have been proposed for POLSAR,
  and demonstrates that almost all of them are based on the likelihood statistical test for complex Wishart distribution.
While an exact statistical distribution is expected to be necessary for a test,
  only an asymptotic distribution is used in the underlying approach \cite{Conradsen_2003_TGRS_4}.

@@

While these models undoubtedly help to further our understandings of the POLSAR data,
  none of the underlying observables have been shown to meet the dual criteria of
  1) resulting in statistically consistent discrimination measures and thus 
  2) being representative of the complex POLSAR data.

@@

The commonly used measure of distance for matrices are either the Euclidean or Manhattan distances, defined respectively as:
\begin{align}
  d(C_x,C_y) &= \sum_{i,j} |\mathbb{R} (C_x - C_y)_{i,j}| + \sum_{i,j} |\mathbb{I} (C_x - C_y)_{i,j}| \\
  d(C_x,C_y) &= \sqrt{\sum_{i,j} |C_x - C_y|_{i,j}^2 }
\end{align}
where $C_{i,j}$ denotes the $(i,j)$ elements of the POLSAR covariance matrix C,
 $||$ denotes absolute values
and $\mathbb{R},\mathbb{I}$ denote the real and imaginary parts respectively.
However, in the context of POLSAR, % covariance matrix
  these dissimilarity measures are not widely used 
  mainly due to the multiplicative nature of the noisy data.

In the field of POLSAR, the Wishart distance is probably the most widely used, as part of the well-known Wishart classifier \cite{Lee_1999_TGRS}.
It is defined \cite{Lee_1994_IJRS_2299} as:
\begin{equation}
  d(C_x,C_y) = \ln|C_y| + tr(C_xC_y^{-1})
\end{equation},
where $tr(C)$ denotes the trace of the matrix C. 
As a measure of distance, its main disadvantage is that $d(C_y,C_y) = \ln|C_y| \neq 0$.

@@

To detect changes, a test statistic is developed for this discrimination measure.
This means a distribution is to be derived for the dissimilarity measure.
However, in the original work \cite{Conradsen_2003_TGRS_4}, only an asymptotic distribution was used.
This paper proposes a statistical model for the determinant of the POLSAR covariance matrix $|C_v|$
  which is capable of providing an exact distribution for the test.

@@

In this paper, the POLSAR scattering vector is denoted as $s$.
For partial polarimetric SAR (single polarization in transmit and dual polarization in receipt),
  the vector is two-dimensional ($d=2$) and is normally written as: 
\begin{equation}
s_{part}=\begin{bmatrix}
S_h\\ 
S_v
\end{bmatrix}
\end{equation}
For full and monostatic POLSAR data,
  the vector is three-dimensional ($d=3$) and is presented as:
\begin{equation}
s_{full}=\begin{bmatrix}
S_{hh}\\
\sqrt{2}S_{hv}\\
S_{vv}
\end{bmatrix}
\end{equation}

@@

The sample POLSAR covariance matrix is formed as the mean of Hermitian outer product of independent single-look scattering vectors,
\begin{equation}
  C_v = \langle ss^{*T} \rangle = \frac{1}{L} \sum^L_{i=1}s_is_i^{*T}
\end{equation}
where $L$ is the number of looks
  and $s_i$ denotes the single-look scattering vector,
  which equals $s_{part}$ for the case of partial POLSAR or
  $s_{full}$ for the case of full polarimetry.

@@

Similar to the way intensity-ratio is proposed as the discrimination measure for the multiplicative and heteroskedastic SAR intensity \cite{Rignot_1993_TGRS_896},
in this paper, the determinant-ratio and the change-ratio are also proposed as a discrimination measure for the POLSAR data.
  which is shown above to also suffer from the multiplicative and heteroskedastic phenomena.
%Similar to the way dispersion and contrast is defined for one-dimensional SAR \cite{Le_2010_ACRS},
%  this section introduces the consistent sense of distance for POLSAR.

@@

Since each elementary component follows fixed distributions (i.e. $\chi^2(2L)$),
  it is natural that this variable also follows fixed distributions.
Moreover, it is independent of the underlying signal $\Sigma_v$,
  indicating its statistically consistent properties and hence its applicability as a POLSAR discrimination measure.
This claim is extrapolated from the widely-used intensity-ratio as a SAR discrimination measure
  and will be further discussed in the next sections.
%is further strengthened in the next section which shows that the generic models for POLSAR presented in this paper also include the commonly-used models for SAR as its special 1-dimensional case. 
%This result shows how
%In short, these random variables are shown to follow consistent and fixed distributions,
%  regardless of the underlying signal $\Sigma_v$.

@@

Since the PDF of chi-squared distribution can be written as:
\begin{align*}
\chi(2L) \sim pdf \left[ \frac{\chi^{L-1}e^{-\chi/2}}{2^L\Gamma(L)} \right]
\end{align*}
Applying variable change theorem into the above equations results in:
\begin{align*}
  |C_v| &\sim  pdf \left[ \frac{L^L x^{L-1} e^{-Lx/|\Sigma_v|}}{\Gamma(L) |\Sigma_v|^L} \right] \\ %\label{eqn:determinant_distribution} %\\
  \mathbb{R}_{\Sigma} = \frac{|C_v|}{|\Sigma_v|} &\sim pdf \left[ \frac{ L^{L} x^{L-1} e^{-Lx}}{ \Gamma(L)} \right] \\ %\label{eqn:polsar_ratio_det_cov_dist} \\
  \mathbb{R}_{C} = \frac{C_1}{C_2} &\sim pdf \left[ \frac{\Gamma(2L-1) x^{L-1}}{\Gamma^2(L-1) (1+x)^{2L}} \right]
\end{align*}
These equations match exactly with the following traditional model for multi-look SAR intensity:
  \begin{eqnarray}
I &\sim& pdf \left[ \frac{L^L x^{L-1} e^{-Lx/\bar{I}}}{\Gamma(L) \bar{I}^L} \right] \\
\mathbb{R}_{\bar{I}} = \frac{I}{\bar{I}} &\sim& pdf \left[ \frac{ L^{L} x^{L-1} e^{-Lx}}{ \Gamma(L)} \label{eqn:multi_look_SAR_ratio_dist} \right] \\
  \mathbb{R}_{I} = \frac{I_1}{I_2} &\sim& pdf \left[ \frac{\Gamma(2L-1) x^{L-1}}{\Gamma^2(L-1) (1+x)^{2L}} \right]
  \end{eqnarray}
considering that $|C_v| \mapsto I$ and $|\Sigma_v| \mapsto \bar{I}$ as multi-dimensional POLSAR collapses into single-dimensional SAR.

@@

As for the first matter, in view of the models given in Eqn \ref{eqn:determinant_distribution},
  the likelihood test statistics presented in \cite{Conradsen_2003_TGRS_4} and rewritten in Eqns \ref{eqn:ori_likelyhood_stats} \& \ref{eqn:log_likelyhood_stats}
can be expressed as:
\begin{align*}
  \ln{Q} &\sim  k + L_x \Lambda^d_{L_x} + L_y \Lambda^d_{L_y} - (L_x + L_y) \Lambda^d_{(L_x + L_y)} \\
  Q &\sim e^k \frac{(\chi^d_{L_x})^{L_x} \cdot (\chi^d_{L_y})^{L_y}}{(\chi^d_{L_x + L_y})^{L_x + L_y}}   
\end{align*}
where $k = d \left[ (L_x + L_y) \ln(L_x + L_y) - L_x \ln{L_x} - L_y \ln{L_y} \right]$.
This, in essence, derives an exact statistical distribution for the likelihood test statistics,
  as opposed to the asymptotic distribution derived in \cite{{Conradsen_2003_TGRS_4}}.

@@

Since it has been shown in Section \ref{sec:sar_special_case_of_polsar} that the model for the case of $d=1$ matches exactly with the traditional model for SAR intensity,
  the validation of the proposed $d=1$ model is quite straightforward. %only a simple validation is presented for this case.  
Fig. \ref{fig:verify_POLSAR_model_1D} presents the results of a test performed for the stated purpose.
In the test, the intensity of single-channel SAR data (HH) for sample homogeneous areas in both datasets is extracted.
The histograms for both the intensity and intensity ratio %the log-distance and contrast
  are then plotted against the theoretical PDF given above.
%Also presented are the plots of the same test applied on a homogeneous patch inside the nine-look processed RADARSAT2 image.
The plots are obtained having set ENL to an estimated value obtained using the technique presented in \cite{Anfinsen_2009_TGRS_3795}.
In all cases, the good visual match between the actual data and model distribution tends to validate the proposed model. %providing us a simple validation of the result.

@@

\plotWithLegend{../images/verify_change_ratio_model_on_RADARSAT2_1d.eps}{change-ratio}
%		 \epsfxsize=1.5in
%		 \epsfysize=1.5in
%		 \epsffile{../images/verify_change_ratio_model_on_RADARSAT2_1d.eps} 	
		 \label{RADARSAT2_1D_det_ratio}
	}
\end{tabular}
\caption{\scriptsize{The specific d=1 models are validated for both RADARSAT2 and AIRSAR datasets.
Each histogram plots the signal under investigation (along the x-axis) against its probabilistic distribution across the patch (on the y-axis),
with p-value being the significant level of the Kolmogorov-Smirnov goodness-of-fit test between the two histograms}}
\label{fig:verify_POLSAR_model_1D}
\end{figure}

@@

The remainder of this section now focuses on validating the models for both partial ($d=2$) and full ($d=3$) POLSAR.
Similar tests %like the one described for SAR data 
  are carried out
  for both types of polarimetric SAR
  on both AIRSAR and RADARSAT2 datasets.
The look-number is estimated for each type of data and for each image using the technique described in \cite{Anfinsen_2009_TGRS_3795}. %and is also noted in the plots.
The practical and predicted distributions are again plotted to determine whether their visual match is close in each case. Fig. \ref{fig:verify_det_ratio_model_2D} shows the $d=2$ plots for determinant, determinant ratio and change ratio for both datasets. It is very clear that the model histogram matches the data well, in fact appearing similar to a smoothed data response. Evidently the match is good for $d=2$.

Similarly, Fig. \ref{fig:verify_det_ratio_model_3D} explores the $d=3$ case for the same data, same model parameters.
Again, although the histogram is much tighter, the match is visually obvious in all cases.
This qualitative observation is further supported by high significance level (or p-value) of the standard Kolmogorov-Smirnov goodness-of-fit test between curves, noted under each graph.
While a good p-value does not ensure automatic acceptance of the model,
  consistently high p-values lend high level of confidence to the proposed model\cite{Kolmogoroff_1941_AMS_461}.

>>>>

%\author{Thanh-Hai Le, Ian McLoughlin}
\author{Thanh-Hai~Le,
        Ian~Vince~McLoughlin, 
	and Chan-Hua~Vun%
\thanks{Thanh-Hai~Le and Chan-Hua~Vun are with School of Computer Engineering, 
Nanyang Technological University, Singapore. Ian~McLoughlin is with the School of Information Science and Technology,
University of Science and Technology of China.
}% <-this % stops a space
%\thanks{The authors wish to thank Dr. Ken-Yoong Lee and Dr. Timo Brestchneider of EADS Innovation-Works Singapore for 
%	providing us the RADAR-SAT2 imagery used in this paper. }% <-this % stops a space
\thanks{Manuscript received March, 2014; revised March, 2014.}}

\markboth{IEEE Journal on Selected Topics in Applied Earth Observation and Remote Sensing,~Vol.~x, No.~x, x~2014}%
{ Le \MakeLowercase{\textit{et al.}}:  Scalar and Representative Observables, and Their Associated Statistical Model, for Polarimetric SAR Data}

@@

%How do others solve this problem?
%The complex POLSAR data have been statistically modelled as following the complex Wishart distribution,
%  which apparently is multidimensional, complex and thus not very intuitive.
%There has been a few scalar observables with accompanying statistical models proposal for POLSAR
A few scalar POLSAR observables with accompanying statistical models have been proposed \cite{Conradsen_2003_TGRS_4, Alberga_2008_IJRS_4129, Joughin_1994_TGRS_562, Lee_1994_TGRS_1017, Touzi_1996_TGRS_519, Lopez-Martinez_2003_TGRS_2232, Erten_2012_Sensors_2766},
  but none of them is able to provide widely-accepted scalar discrimination measures.
%  but none of them has led to scalar discrimination measures proposal.
As such no observable can be considered as being highly representative of this multi-dimensional data,
  which severely limits their applicability in practical data processing applications.
Alternatively, a few popular POLSAR discrimination measures have been proposed~\cite{Lee_1994_IJRS_2299, Anfinsen_2007_ESA_POLINSAR, Kersten_2005_TGRS_519, Lee_2011_IGARSS_3740, Cao_2007_TGRS_3454},
%From an alternative approach, a few POLSAR discrimination measures have been proposed,
  but all of them are based on likelihood ratio statistics (see Subsection \ref{sec:lit_measures}).
This statistical test should ideally be based on an exact and consistent distribution
   but so far only asymptotic distributions have been demonstrated \cite{Conradsen_2003_TGRS_4}.  
%This statistical test should be based on an exact and consistent statistical distribution
%  but so far has only shown to be based on an asymptotic distribution.

%Research Objective: Give your precise problem statement
This paper presents scalar and representative observables,
   and an associated generic statistical model,
   that describes multi-dimensional POLSAR data to provide a consistent foundation for the derivation of discrimination measures.   
%This article hence presents a scalar and representative observable, and its associated generic statistical model to describe multi-dimensional POLSAR data and provide a consistent foundation for the derivation of POLSAR discrimination measures.
%Now State your approach
The development is based on the common assumption that POLSAR data follows the complex Wishart distribution.  
Subsequently, the generic statistical model for the covariance matrix determinant is presented as being just a scalar projection of the multidimensional data.
This model is then used to derive several scalar and consistent statistical distributions suggesting that their associated observables are capable of being used as discrimination measures for POLSAR.
%This study started off with the observation that the determinant of the covariance matrix is widely used in POLSAR discrimination measures.
%However, our survey of existing scalar models for POLSAR indicates that no statistical model for this observable has been published so far.
%Consequently, in this article, the statistical behaviour of the determinant of POLSAR covariance matrix is further investigated.
%Prof Vuns: comment: the article provides the result, not the journey!
%POLSAR data have been statistically modelled as following the complex Wishart distribution.
%Consequently, the generic statistical model for the covariance matrix determinant is presented as being just a scalar projection of the multidimensional data.
%This model is then used to derive several scalar and consistent statistical descriptions suggesting that their associated observables are capable of being used as discrimination measures for POLSAR data.
%
%***IVM: the following sentence doesn't make sense, also, I don't think it's good to talk about 'models of a model'.  Maybe 'model variants' instead?
%%%***IVM***The specific two and three dimensional models of these models, which %also include the derived discrimination measures, are originally designed against both partial and full polarimetric SAR data.

This paper also shows that %It is also shown that
   the specific one dimensional (1-D) version of the proposed  model 
   matches %perfectly with 
   the traditional statistical model used for SAR intensity.
This effectively incorporates the existing SAR theory under the umbrella of the proposed scalar approach for multi dimensional POLSAR.
The different discrimination proposals for SAR and POLSAR are reviewed in light of this and the  proposal in this paper is shown to provide a strong, unifying and consistent foundation. % relating them all together.
The applicability of these theoretical models will be illustrated by experiments where the specific 1, 2 and 3-D versions of the proposed models are validated against real-life captured data.    
The other beneficial implication of the proposed approach will also be demonstrated and discussed,
  when the proposed model is applied, as an example illustrating their effectiveness, to the evaluation of POLSAR speckle filters.

@@

%Article Organization: Show that you deliver on the objectives and hence solve the stated problem
The remainder of this article is organized as follows.
After the next section reviews existing discrimination measures as well as scalar observables models for POLSAR, Section \ref{sec:theoretical_model}
  derives the generic statistical model for the determinant of the POLSAR covariance matrix
  and proposes several new discrimination measures for the multi-dimensional data.
Section \ref{sec:sar_special_case_of_polsar} then illustrates the correspondence between the specific one dimensional model, and traditional models, for SAR intensity. 
Following this, Section \ref{sec:link_sar_polsar} links the disparate discrimination measure proposals for traditional SAR with more recent POLSAR data.
%Section \ref{sec:polsar_models_validation} then validates the specific one, two and three dimensional versions of the proposed model against real-life captured data.
The applicability of these models against practical data is illustrated in Section \ref{sec:polsar_models_validation}.
%Section \ref{sec:discussion} presents a high-level discussion of the results presented before
Section \ref{sec:application} then presents a simple demonstration where an existing technique of SAR processing is extended for POLSAR data before
  Section \ref{sec:conclusion} finally concludes the paper.

@@

%This section outlines related publication.
Section \ref{sec:lit_models} reviews various statistical models used for different scalar observables for POLSAR,
  where it is shown that none are able to provide statistically consistent discrimination measures. 
Section \ref{sec:lit_measures} further strengthens those findings by discussing discrimination measures that have been proposed for POLSAR,
  and demonstrates that almost all of them are based on the likelihood statistical test for complex Wishart distribution.
While an exact statistical distribution is expected to be necessary for a test,
  only an asymptotic distribution is used in their underlying foundation \cite{Conradsen_2003_TGRS_4}.

@@

While these models undoubtedly help to further our understanding of POLSAR data,
  none of the underlying observables have been shown to meet the dual criteria of
  (i) resulting in statistically consistent discrimination measures and thus 
  (ii) being representative of the complex POLSAR data.

@@

Commonly used measures of dissimilarity for matrices are either the Euclidean or Manhattan distances, defined respectively as:
\begin{align}
  d(C_x,C_y) &= \sum_{i,j} |\mathbb{R} (C_x - C_y)_{i,j}| + \sum_{i,j} |\mathbb{I} (C_x - C_y)_{i,j}| \\
  d(C_x,C_y) &= \sqrt{\sum_{i,j} |C_x - C_y|_{i,j}^2 }
\end{align}
where $C_{i,j}$ denotes the $(i,j)$ elements of the POLSAR covariance matrix C,
 $||$ denotes absolute values
and $\mathbb{R},\mathbb{I}$ denote the real and imaginary parts respectively.
However, in the context of POLSAR, % covariance matrix
  these dissimilarity measures are not widely used 
  mainly due to the multiplicative nature of the noisy data.

In the field of POLSAR, the Wishart distance is probably the most widely used, as part of the well-known Wishart classifier \cite{Lee_1999_TGRS}.
It is defined \cite{Lee_1994_IJRS_2299} as:
\begin{equation}
  d(C_x,C_y) = \ln|C_y| + tr(C_xC_y^{-1})
\end{equation}
where $tr(C)$ denotes the trace of the matrix C. 
As a measure of distance, its main disadvantage is that $d(C_y,C_y) = \ln|C_y| \neq 0$.

@@

To detect changes, a test statistic is developed for this discrimination measure.
This means a distribution is to be derived for the dissimilarity measure.
However, in the original work \cite{Conradsen_2003_TGRS_4}, only an \textit{asymptotic} distribution was used.
By contrast, this paper will propose a statistical model for the determinant of the POLSAR covariance matrix $|C_v|$
  which is capable of providing an \textit{exact} distribution for the test.

@@

Note that the POLSAR scattering vector is denoted as $s$ in this paper.
For partial polarimetric SAR (single polarization in transmit and dual polarization in receive),
  the vector is two-dimensional ($d=2$) and is normally written as: 
\begin{equation}
s_{part}=\begin{bmatrix}
S_h\\ 
S_v
\end{bmatrix}
\end{equation}
For full and monostatic POLSAR data,
  the vector is three-dimensional ($d=3$) and is presented as:
\begin{equation}
s_{full}=\begin{bmatrix}
S_{hh}\\
\sqrt{2}S_{hv}\\
S_{vv}
\end{bmatrix}
\end{equation}

@@

The sample POLSAR covariance matrix is formed as the mean of the Hermitian outer product of independent single-look scattering vectors,
\begin{equation}
  C_v = \langle ss^{*T} \rangle = \frac{1}{L} \sum^L_{i=1}s_is_i^{*T}
\end{equation}
where $L$ is the number of looks
  and $s_i$ denotes the single-look scattering vector,
  which equals $s_{part}$ for the case of partial POLSAR or
  $s_{full}$ for the case of full polarimetry.

@@

Similar to the way intensity-ratio is proposed as the discrimination measure for the multiplicative and heteroskedastic SAR intensity \cite{Rignot_1993_TGRS_896},
in this paper, the determinant-ratio and the change-ratio are also proposed as a discrimination measure for the POLSAR data:
  Which is shown above to also suffer from the multiplicative and heteroskedastic phenomena.
%Similar to the way dispersion and contrast is defined for one-dimensional SAR \cite{Le_2010_ACRS},
%  this section introduces the consistent sense of distance for POLSAR.

@@

Since the elementary components follow fixed distributions (i.e. $\chi^2(2L)$),
  it is natural that this variable also follows a fixed distribution.
Moreover, it is independent of the underlying signal $\Sigma_v$,
  indicating its statistically consistent properties and hence its applicability as a POLSAR discrimination measure.
This claim is extrapolated from the widely-used intensity-ratio as a SAR discrimination measure
  and will be further discussed in the next sections.
%is further strengthened in the next section which shows that the generic models for POLSAR presented in this paper also include the commonly-used models for SAR as its special 1-dimensional case. 
%This result shows how
%In short, these random variables are shown to follow consistent and fixed distributions,
%  regardless of the underlying signal $\Sigma_v$.

@@

Since the PDF of chi-squared distribution can be written as:
\begin{align*}
\chi(2L) \sim pdf \left[ \frac{\chi^{L-1}e^{-\chi/2}}{2^L\Gamma(L)} \right]
\end{align*}
Applying the variable change theorem into the above equations results in:
\begin{align*}
  |C_v| &\sim  pdf \left[ \frac{L^L x^{L-1} e^{-Lx/|\Sigma_v|}}{\Gamma(L) |\Sigma_v|^L} \right] \\ %\label{eqn:determinant_distribution} %\\
  \mathbb{R}_{\Sigma} = \frac{|C_v|}{|\Sigma_v|} &\sim pdf \left[ \frac{ L^{L} x^{L-1} e^{-Lx}}{ \Gamma(L)} \right] \\ %\label{eqn:polsar_ratio_det_cov_dist} \\
  \mathbb{R}_{C} = \frac{C_1}{C_2} &\sim pdf \left[ \frac{\Gamma(2L-1) x^{L-1}}{\Gamma^2(L-1) (1+x)^{2L}} \right]
\end{align*}
These equations match exactly with the following traditional model for multi-look SAR intensity:
  \begin{eqnarray}
I &\sim& pdf \left[ \frac{L^L x^{L-1} e^{-Lx/\bar{I}}}{\Gamma(L) \bar{I}^L} \right] \\
\mathbb{R}_{\bar{I}} = \frac{I}{\bar{I}} &\sim& pdf \left[ \frac{ L^{L} x^{L-1} e^{-Lx}}{ \Gamma(L)} \label{eqn:multi_look_SAR_ratio_dist} \right] \\
  \mathbb{R}_{I} = \frac{I_1}{I_2} &\sim& pdf \left[ \frac{\Gamma(2L-1) x^{L-1}}{\Gamma^2(L-1) (1+x)^{2L}} \right]
  \end{eqnarray}
considering that $|C_v| \mapsto I$ and $|\Sigma_v| \mapsto \bar{I}$ when multi-dimensional POLSAR collapses into single-dimensional SAR.

@@

As for the first matter, in view of the models given in Eqn \ref{eqn:determinant_distribution},
  the likelihood test statistics presented in \cite{Conradsen_2003_TGRS_4} and rewritten in Eqns \ref{eqn:ori_likelyhood_stats} \& \ref{eqn:log_likelyhood_stats}
can be expressed as:
\begin{align*}
  \ln{Q} &\sim  k + L_x \Lambda^d_{L_x} + L_y \Lambda^d_{L_y} - (L_x + L_y) \Lambda^d_{(L_x + L_y)} \\
  Q &\sim e^k \frac{(\chi^d_{L_x})^{L_x} \cdot (\chi^d_{L_y})^{L_y}}{(\chi^d_{L_x + L_y})^{L_x + L_y}}   
\end{align*}
where $k = d \left[ (L_x + L_y) \ln(L_x + L_y) - L_x \ln{L_x} - L_y \ln{L_y} \right]$.
This, in essence, derives an exact statistical distribution for the likelihood test statistics (as opposed to an asymptotic distribution \cite{{Conradsen_2003_TGRS_4}}).

@@

Since it has been shown in Section \ref{sec:sar_special_case_of_polsar} that the model for the case of $d=1$ matches exactly with the traditional model for SAR intensity,
  the validation of the proposed $d=1$ model is quite straightforward. %only a simple validation is presented for this case.  
Fig. \ref{fig:verify_POLSAR_model_1D} presents the results of a test performed for the stated purpose.
In the test, the intensity of single-channel SAR data (HH) for sample homogeneous areas in both datasets is extracted.
The histograms for both the intensity and intensity ratio %the log-distance and contrast
  are then plotted against the theoretical PDF given above.
%Also presented are the plots of the same test applied on a homogeneous patch inside the nine-look processed RADARSAT2 image.
The plots are obtained having set ENL to an estimated value obtained using the technique presented in \cite{Anfinsen_2009_TGRS_3795}.
In all cases, the good match between the actual data and model distribution lends support to the proposed model. %providing us a simple validation of the result.

@@

\plotWithLegend{../images/verify_change_ratio_model_on_RADARSAT2_1d.eps}{change-ratio}
%		 \epsfxsize=1.5in
%		 \epsfysize=1.5in
%		 \epsffile{../images/verify_change_ratio_model_on_RADARSAT2_1d.eps} 	
		 \label{RADARSAT2_1D_det_ratio}
	}
\end{tabular}
\caption{\scriptsize{The specific d=1 models are validated for both RADARSAT2 and AIRSAR datasets.
Each histogram plots the signal under investigation (along the x-axis) against its probabilistic distribution across the patch (on the y-axis),
with p-value being the significance level of the Kolmogorov-Smirnov goodness-of-fit test between the two histograms}}
\label{fig:verify_POLSAR_model_1D}
\end{figure}

@@

The remainder of this section now focuses on validating the models for both partial ($d=2$) and full ($d=3$) POLSAR.
Similar tests %like the one described for SAR data 
  are carried out
  for both types of polarimetric SAR
  on both AIRSAR and RADARSAT2 datasets.
The look-number is estimated for each type of data and for each image~\cite{Anfinsen_2009_TGRS_3795}. %and is also noted in the plots.
The practical and predicted distributions are again plotted to determine whether their visual match is close in each case. Fig. \ref{fig:verify_det_ratio_model_2D} shows the $d=2$ plots for determinant, determinant ratio and change ratio for both datasets. It is very clear that the model histogram matches the data well, in fact appearing similar to a smoothed data response. Evidently the match is good for $d=2$.

Similarly, Fig. \ref{fig:verify_det_ratio_model_3D} explores the $d=3$ case for the same data, same model parameters.
Again, although the histogram is much tighter, the match is visually obvious in all cases.
This qualitative observation is further supported by high significance level (or p-value) in the standard Kolmogorov-Smirnov goodness-of-fit test between curves~\cite{Kolmogoroff_1941_AMS_461}, noted under each graph.
While a good p-value may not lead to automatic acceptance of the model,
  consistently high p-values lend a high level of confidence to the proposal.
